\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-10-28 paper.tex
update beta simulation formula
rewrite lemma 1 proof

Now, concerning the simulation of the longitudinal data, the idea is to sample from multivariate normal distributions. 
We use linear time-varying features such that
\[ Y_i^l(t) = \sum_{k=0}^{K-1} \ind{\{G_i=k\}} \big( (1, t)^\top \beta_k^l + (1, t)^\top b_i^l + \epsilon_i^l(t) \big) \]
where $t \geq 0$, the error term $\epsilon_i^l(t) \sim \cN(0, \sigma_l^2)$, the global variance-covariance matrix for the random effects components~\eqref{eq:random-effect-law} is such that $D = \bSigma_2(\rho_2)$, a $(r \times r)$ Toeplitz covariance matrix with correlation $\rho_2 \in (0, 1)$, and the fixed effect parameters are generated according to 
\[ \beta_k^l \sim \cN\Big( 
\mu_k, 
\begin{bmatrix}
  \rho_3 & 0\\
  0 & \rho_3
\end{bmatrix} 
\Big) \]
for $k \in \{0, 1\}$ and with correlation $\rho_3 \in (0, 1)$, so that $\beta_k^l$ coefficients are both positive for high-risk subjects ($G_i=1$), which favours increasing trajectories with positive values, and negative otherwise ($G_i=0$), which favours decreasing trajectories with negative values. This choice makes sens regarding that for simplicity, we take joint association parameters being all positive, see~\eqref{eq:gamma-sparse}.
We simulate measurement times using a regular grid, and an example of simulated longitudinal features for an individual is displayed in Figure~\ref{fig:example-data} below.

@@

\texttt{TODO Simon : rewrite the proof properly}

The proximal operator for any function $f : R^p \to R$ is defined by
\begin{equation}
    \label{eq:prox-def}
    \text{prox}_{f}(u) = \argmin_{x \in \R^p} \{\frac{1}{2}\norm{x - u}_2^2 + f(x)\}.
\end{equation}
\begin{lemma} 
The following holds
\begin{equation*}
    \text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L  \norm{\cdot}_{2,l}}
    = \text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}_{2,l}} \circ \text{prox}_{\lambda_1 \norm{\cdot}_1}
\end{equation*}
where $prox_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}_2}$ is the proximal operator for the sparse group lasso regulariser, $prox_{\lambda_2 \sum_{l=1}^L \norm{\cdot}_2}$ is the proximal operator for the group lasso regulariser and $prox_{\lambda_1 \norm{\cdot}_1}$ is the proximal operator for the lasso regulariser. $\lambda_1 > 0$, $\lambda_2 > 0$ are regularization parameters, $\norm{\cdot}_{2,l}$ denotes the Euclidean norm on features in l-th group.
\end{lemma}

\begin{proof}
Denote 
\begin{equation*}
    k(x) =  \frac{1}{2}\norm{ x - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)}_2^2 + \lambda_2 \sum_{l=1}^L \norm{x}_{2,l}
\end{equation*}
then we have the proximal operator (see~\eqref{eq:prox-def}) for the group lasso regulariser with $u = \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)$ is given by
\begin{equation*}
    \text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}}\big(\text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)\big) = \argmin_x k(x).
\end{equation*}
The sub-differential \texttt{[REF]} of $k(\cdot)$ at $x$ is given by
\begin{equation*}
    \partial k(x) = x - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) + \partial( \lambda_2 \sum_{l=1}^L \norm{x}_{2,l} )
\end{equation*}
It is clear that $k(\cdot)$ is convex function. Denote the unique minimizer of $k(\cdot)$ is $x^\star$. The sufficient and necessary condition for the optimality of $x^\star$ is 
\begin{equation}
    \label{eq:opt_gl}
    0 \in x^* - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2,l} ).
\end{equation}
\texttt{[REF]}.
Denote $y = \text{prox}_{\lambda_1\norm{\cdot}_1}(\beta)$. From Lemma 1 in ~\citet{yuan2011efficient}, one has that $\text{SGN}(y) \subseteq \text{SGN}(x^*)$\\
where \[\text{SGN}(x) = 
\left\{
  \begin{array}{ll}
    \{1\} &\text{if } x > 0, \\ 
    $[-1, 1]$ &\text{if } x = 0, \\ 
    \{-1\} &\text{if } x < 0. 
  \end{array}
\right. \]
which is the sub-differential of the l1 norm. \texttt{[REF]} \\
\noindent Denote 
\begin{equation*}
    h(x) =  \frac{1}{2}\norm{x - \beta}_2^2 + \lambda_1 \norm{x}_1 + \lambda_2 \sum_{l=1}^L \norm{x}_{2,l}
\end{equation*}
then we have the  proximal  operator  for  the  sparse  group  lasso  regulariser is given by
\begin{equation*}
    \text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}}(\beta) = \argmin_{x} \{ h(x)\}.
\end{equation*}
Denote the sub-differential of $h(\cdot)$ at $x$ is given by
\begin{align*}
    \partial(h(x)) &= x - \beta + \partial(\lambda_1 \norm{x}_1) + \partial( \lambda_2 \sum_{l=1}^L \norm{x}_{2, l} ) \\
    &= x - \beta + \lambda_1 \text{SGN}(x) + \partial( \lambda_2 \sum_{l=1}^L \norm{x}_{2, l} )
\end{align*}

\noindent We need to prove that $0 \in \partial\big(h(x^\star)\big)$. We can easily prove that 
\begin{equation*}
    y = \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) 
    = \argmin_{u} \{\norm{u - \beta}_2^2 + \lambda_1 \norm{u}_1\} 
    = \beta - \lambda_1 \text{SGN}(y) \in \beta - \lambda_1 \text{SGN}(x^*).
\end{equation*}
Hence
\begin{equation}
    \label{eq:opt_gl_in_sgl}
    x^\star - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2, l} ) \in x^\star- \beta + \lambda_1 \text{SGN}(x^\star) + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2, l} )
\end{equation}
It follows from ~\eqref{eq:opt_gl}- ~\eqref{eq:opt_gl_in_sgl} that $0 \in \partial(h(x^\star))$. Hence
\begin{equation*}
    \text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}} = \text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}} \circ \text{prox}_{\lambda_1 \norm{\cdot}_1}
\end{equation*}
\end{proof}

>>>>

Now, concerning the simulation of the longitudinal data, the idea is to sample from multivariate normal distributions. 
We use linear time-varying features such that
\[ Y_i^l(t) = \sum_{k=0}^{K-1} \ind{\{G_i=k\}} \big( (1, t)^\top \beta_k^l + (1, t)^\top b_i^l + \epsilon_i^l(t) \big) \]
where $t \geq 0$, the error term $\epsilon_i^l(t) \sim \cN(0, \sigma_l^2)$, the global variance-covariance matrix for the random effects components~\eqref{eq:random-effect-law} is such that $D = \bSigma_2(\rho_2)$, a $(r \times r)$ Toeplitz covariance matrix with correlation $\rho_2 \in (0, 1)$, and the fixed effect parameters are generated according to 
\[ \beta_k^l \sim \ind{\{l \in \cS_k \}} \cN\Big( 
\mu_k, 
\begin{bmatrix}
  \rho_3 & 0\\
  0 & \rho_3
\end{bmatrix} 
\Big)
+ \ind{\{l \notin \cS_k \}} \cN\Big( 
\begin{pmatrix}
  0 \\ 0
\end{pmatrix}
, 
\begin{bmatrix}
  \rho_4 & 0\\
  0 & \rho_4
\end{bmatrix} 
\Big) \]
for $k \in \{0, 1\}$ and with correlation $\rho_3 \in (0, 1)$, so that $\beta_k^l$ coefficients are both positive for high-risk subjects ($G_i=1$), which favours increasing trajectories with positive values, and negative otherwise ($G_i=0$), which favours decreasing trajectories with negative values. This choice makes sens regarding that for simplicity, we take joint association parameters being all positive, see~\eqref{eq:gamma-sparse}.
We simulate measurement times using a regular grid, and an example of simulated longitudinal features for an individual is displayed in Figure~\ref{fig:example-data} below.

@@

The proximal operator~\citep{moreau1962fonctions} for any function $f : \R^p \to \R$ is defined by
\begin{equation}
    \label{eq:prox-def}
    \text{prox}_{f}(u) = \argmin_{x \in \R^p} \big\{\frac{1}{2}\norm{x - u}_2^2 + f(x)\big\}.
\end{equation}
To prove Lemma~\ref{lemma:prox-sgl1}, we need to show that
\[\text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L  \norm{\cdot}_{2,l}} = \text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}_{2,l}} \circ \text{prox}_{\lambda_1 \norm{\cdot}_1}\]
where $\text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}_{2,l}}$ is the proximal operator for the sparse group lasso regularizer, $\text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}_{2,l}}$ is the proximal operator for the group lasso regularizer and $\text{prox}_{\lambda_1 \norm{\cdot}_1}$ is the proximal operator for the lasso regularizer. $\lambda_1 > 0$ and $\lambda_2 > 0$ are regularization parameters, and $\norm{\cdot}_{2,l}$ denotes the Euclidean norm on features in the $l$-th group.

\begin{proof}
Fix any $\beta \in \R^p$ and denote 
\begin{equation*}
    k_\beta(x) =  \frac{1}{2}\norm{ x - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)}_2^2 + \lambda_2 \sum_{l=1}^L \norm{x}_{2,l}
\end{equation*}
for all $x \in \R^p$. Given that $\big(\text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)\big)_j=\max(0, 1-\lambda_1/|\beta_j|)\beta_j$~\citep{bach2011optimization}, $k_\beta$ is clearly strictly convex. Denoting $x^\star$ its unique minimizer,~\eqref{eq:prox-def} entails 
\[\text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}_{2,l}}\big(\text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)\big) = \argmin_{x \in \R^p} k_\beta(x) = \{x^\star\}.\]
Then, it follows that
\begin{equation}
    \label{eq:opt_gl}
    0 \in \partial k(x^\star) = x^\star - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) + \partial\big( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2,l}\big)
\end{equation}
where $\partial k$ denotes the subgradient~\citep{rockafellar1974} of $k$.
Now denoting
\[h_\beta(x) = \frac{1}{2}\norm{x - \beta}_2^2 + \lambda_1 \norm{x}_1 + \lambda_2 \sum_{l=1}^L \norm{x}_{2,l} \] 
for all $x \in \R^p$ such that
\[\text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}_{2,l}}(\beta) = \argmin_{x\in \R^p} h_\beta(x),\]
it remains to be proved that $0 \in \partial h_\beta(x^\star)$.

Then, it turns out that 
\[ \partial h_\beta(x) = x - \beta + \lambda_1 \text{SGN}(x) + \partial( \lambda_2 \sum_{l=1}^L \norm{x}_{2, l})\]
for all $x \in \R^p$, where SGN($x$) is the subgradient of the $\ell_1$-norm at $x$~\citep{bach2011optimization} defined in a component wise fashion as
\[\text{SGN}(x) = 
\left\{
  \begin{array}{ll}
    \{1\} &\text{if } x > 0, \\ 
    $[-1, 1]$ &\text{if } x = 0, \\ 
    \{-1\} &\text{if } x < 0. 
  \end{array}
\right. \]
Denoting $y = \text{prox}_{\lambda_1\norm{\cdot}_1}(\beta)$, Lemma 1.4) in~\citet{yuan2011efficient} entails $\text{SGN}(y) \subseteq \text{SGN}(x^*)$. Hence,
\[ y = \argmin_{u \in \R^p} \big\{\frac{1}{2} \norm{u - \beta}_2^2 + \lambda_1 \norm{u}_1\big\} 
    = \beta - \lambda_1 \text{SGN}(y) \subseteq \beta - \lambda_1 \text{SGN}(x^*), \]
which gives
\[0 \in x^\star - y + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2, l} ) \subseteq x^\star- \beta + \lambda_1 \text{SGN}(x^\star) + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2, l} ).\]
\end{proof}
